Reading in the ENSG_ID2Name.txt text file (ENSEMBL gene IDs) and corresponding gene symbols. 
```{r}
gene_symbol_mapping <-  read.table("ENSG_ID2Name.txt", header=FALSE, stringsAsFactors = FALSE, row.names = 1)
```

```{r}
raw_count_data <- read.table("final_read_counts.tsv", header =TRUE, stringsAsFactors = TRUE, row.names=1)
```


Removed last 5 rows of data and stored in new data matrix (minus the last 5 rows) in new variable called raw_count_data2. 
```{r}
raw_count_data2 <- head(raw_count_data, -5)

```


Required at least 25% of samples to have count > 25 

```{r}
quant <- apply(raw_count_data2,1,quantile,0.75)
keep <- which((quant >= 25) == 1)
filtered_raw_count_data <- raw_count_data2[keep,]
```

load edgeR
```{r}
library(edgeR)
```

class labels for Dex treated vs untreated.  
```{r}
class <- factor( c( rep("Dex",3), rep("untreated",3) ))
```

Getting the gene names from gene_symbol_mapping for the genes in the filtered_raw_count_data variable
```{r}
genes <- rownames(filtered_raw_count_data)
gene_names <- gene_symbol_mapping[genes,1]
```


Made a DGEList object 

```{r}
tutorial_dgelist <- DGEList(counts = filtered_raw_count_data, genes = gene_names, group = class)
```


Performed EdgeR Normalization on the list using the calcNormFactors() function
```{r}
tutorial_dgelist <- calcNormFactors(tutorial_dgelist)
```


Exploring data: usinb the *plotMDS* function to look at the distances between samples. 

```{r}
plotMDS(tutorial_dgelist)
```

 
Dispersion estimate for DEG analysis
```{r}
tutorial_dgelist <- estimateCommonDisp(tutorial_dgelist, verbose=TRUE) 
tutorial_dgelist <- estimateTagwiseDisp(tutorial_dgelist)  
```


Performed *exact test* ondata to look for differentially expressed genes using the exactTest() function and stored in a new variable called tutorial_et.
```{r}
tutorial_et <- exactTest(tutorial_dgelist, pair= c("Dex", "untreated"))
```

Printed top differentially expressed genes from the exact test using the topTags() function
```{r}
topTags(tutorial_et)
```


Corrected for multiple hypothesis testing using the decideTests() function
 
```{r}
tutorial_de <- decideTests(tutorial_et, method = "seperate", adjust.method = "fdr", p.value < 0.05)
summary(tutorial_de)
```


Output differentially expressed genes in a Matrix of significantly DE genes:
```{r}
tutorial_de_gene_matrix = cbind(
 genes,gene_names,
 sprintf('%0.3f',log10(tutorial_et$table$PValue)),
 sprintf('%0.3f',abs(tutorial_et$table$logFC))
)[as.logical(tutorial_de),] 
colnames(tutorial_de_gene_matrix) = c("Gene", "Gene_Name", "Log10_Pvalue", "Log_fold_change")
```

Ordered by log fold change and filtered for genes that only have a logFC greater or equal to 1.03 same as Himes et al: 
```{r}
o <- order(tutorial_et$table$logFC[as.logical(tutorial_de)],decreasing=TRUE)
tutorial_de_gene_matrix <- tutorial_de_gene_matrix[o,]
tutorial_de_gene_matrix[ ,2][tutorial_de_gene_matrix[ ,4] >= 1.03]
```

