---
title: "Volcano Plot Tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read in the DE_genes.txt file you generated last class
```{r}
DE_genes <- read.table("DE_genes.txt", header= TRUE)
```

#Check the contents of the file using the head command
```{r}
head(DE_genes)
```

#Make a basic volcano plot

plot() is a generic function used for plotting R objects. 
Usage plot(x, y, ...)

Required arguments to supply:
x = Log_fold_change column from DE_genes
y = negative Log10_Pvalue from DE_genes
DE_genes$Log10_Pvalue <- DE_genes$Log10_Pvalue*(-1)

```{r, echo=FALSE}
plot(DE_genes[,4], DE_genes[,3])
```

#Make your volcano plot look nicer

Optional arguments:
pch= tells R what type symbol you want your points to be. Try setting this to 20
xlab= Add a different x axis title. Try changing the x axis title to "Log2 Fold Change"
ylab= Add a different y axis title. Try changing the y axis to "-log10 Pvalue"
main= add a title to your plot. Try adding a title "Tutorial DE Genes Volcano Plot"


```{r, echo=FALSE}
plot(DE_genes[,4], DE_genes[,3], pch=20, xlab= "Log2 Fold Change", ylab= "-log10 Pvalue", main= "Tutorial DE Genes Volcano Plot")
```


#Add colored points to your plot
What if you want to highlight specific points in your plot? For example to highlight points which have a particular log fold change or p value cut off. 

To highlight genes which have an absolute log fold change greater than 1
1) Plot your data using the plot function using the same code as line 44 above
2) Subset your list of DE_genes to select genes which have a absolute logfold change greater than 1. Subset command usage subset(x, subset, ...)
3) Use the points function to color the points from your subsetted list orange. Points command usage points(x, y, ...) 

```{r}
plot(DE_genes[,4], DE_genes[,3], pch=20, xlab= "Log2 Fold Change", ylab= "-log10 Pvalue", main= "Tutorial DE Genes Volcano Plot")
DE_subset <- subset(DE_genes, abs(DE_genes$Log_fold_change) >1)
points(DE_subset[,4],DE_subset[,3] , pch=20, col="orange")

```


