---
title: "DESeq2"
author: "me"
date: "2023-10-04"
output: html_document
---
#Install packages and load libraries


```{r setup, include=FALSE}
install.packages("htmltools")
library(htmltools)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.16")

BiocManager::install("DESeq2")
library( "DESeq2" )

install.packages('tidyverse')
library('tidyverse')

BiocManager::install(ggplot)
library(ggplot2)
```

## insert countData and colData(sample types)



```{r cars}
raw_count_data <- read.table("final_read_counts.tsv", header =TRUE, stringsAsFactors = TRUE, row.names=1)

metdat1 <- read.csv("metdata_2.csv", header = TRUE, sep = "\t")
```

## Required at least 25% of samples to have count > 25 

```{r pressure, echo=FALSE}
quant <- apply(raw_count_data2,1,quantile,0.75)
keep <- which((quant >= 25) == 1)
filtered_raw_count_data <- raw_count_data2[keep,]
```

#Construct DESEQDataSet Object
```{r}
dds2 <- DESeqDataSetFromMatrix(countData=filtered_raw_count_data, 
                               colData=metdat1,
                               design=dex)
```

#Run DESEQ function
```{r}
dds <- DESeq(dds)
```

#View results table and check summary of differential gene expression
```{r}
res <- results(dds)
head(results(dds, tidy=TRUE))

summary(res)
```

#Sorting summary list by p values
```{r}
res <- res[order(res$padj),]
head(res)
```

#Use of plotCounts fxn to compare the normalized counts between treated and control groups for our top 6 genes
```{r}
par(mfrow=c(2,3))

plotCounts(dds, gene="ENSG00000189221", intgroup="dex")
plotCounts(dds, gene="ENSG00000139132", intgroup="dex")
plotCounts(dds, gene="ENSG00000166741", intgroup="dex")
plotCounts(dds, gene="ENSG00000162692", intgroup="dex")
plotCounts(dds, gene="ENSG00000096060", intgroup="dex")
```
#reset par and Make a basic volcano plot
# colored points: blue if padj<0.01, red if log2FC>1 and padj<0.05)
```{r}
par(mfrow=c(1,1))
with(res2, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-3,3)))
with(subset(res2, padj<.01 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(res2, padj<.01 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

#Transform the raw count data
#vst function will perform variance stabilizing transformation
```{r}
vsdata <- vst(dds, blind=FALSE)
```
#using the DESEQ2 plotPCA fxn to see how samples group
```{r}
plotPCA(vsdata, intgroup="dex")
```
#likelihood ratio for multiple hypothesis testing
```{r}
dds <- DESeq(dds, test="LRT", reduced=~1)
res <- results(dds)
```


