---
title: "Heatmap tutorial"
output: html_document
---

---
title: "Heatmaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install RColorBrewer and gplots packages 
gplots is a popular package for visualizing data. 
RColorBrewer is a package which allows you to clustomize your heatmap colors.


#Load edgeR, RColorBrewer and gplots 
```{r, echo=FALSE}
library(edgeR)
library(RColorBrewer)
library(gplots)
```


#Convert raw reads into log counts per million for visualization with a heatmap

Read in your count matrix and remove the last 5 lines
```{r}
raw_count_data <- read.table("gene_read_counts_table_all_final.tsv", header=TRUE, stringsAsFactors=FALSE, row.names=1)
raw_count_data <- head(raw_count_data, -5)
tail(raw_count_data)
```


Require at least 25% of samples to have count > 25 
```{r}
raw_count_data <- head(raw_count_data, -5)
quant <- apply(raw_count_data,1,quantile,0.75)
keep <- which((quant >= 25) == 1)
raw_count_data <- raw_count_data[keep,]
head(raw_count_data)
```

load edgeR
```{r}
library(edgeR)
```


Make class labels (normal vs tumor). This is your two experimental conditions you want to compare, in our case it is normal versus tumor.
```{r}
class <- factor( c( rep("normal",3), rep("tumor",3) ))
```


Make a DGEList object
```{r}
tutorial_dgelist <- DGEList(counts=raw_count_data, group=class)
```


Convert read counts into log2-counts-per-million (logCPM). The edgeR logcmp function does this for a DGElist. 
The prior.counts argument adds 2 to each observation to prevent taking the log of 0. 
```{r}
logcpm <- cpm(tutorial_dgelist, prior.count=2, log=TRUE)
head(logcpm)
```


#Generate a heatmap

##Get some pretty colors
```{r}
mypalette <- brewer.pal(11,"RdYlBu")
morecols <- colorRampPalette(mypalette)
```


##Set up colour vector for cell type variable
```{r}
col.cell <- c("yellow","orange")[class]
```


##Plot the heatmap

```{r}
heatmap.2(logcpm, col=rev(morecols(50)), trace="none", main="Tumor vs. Normal", ColSideColors=col.cell, scale="row", srtCol=45, cexCol=0.75, cexRow=0.5)
```


##Save your heatmap as a png file
```{r, echo=FALSE}
png(file="Tumor_vs_Normal_heatmap.png")
heatmap.2(logcpm,col=rev(morecols(50)),trace="none", main="Tumor vs. Normal", ColSideColors=col.cell, scale="row", srtCol=45, cexCol=0.75, cexRow=0.5)
dev.off()
```



#Modifying your heatmap

Change the color scheme of your heatmap. Use ?RColorBrewer to see available colour schemes. Replot your heatmap with a new color scheme (use diverging palettes).

```{r}
mypalette <- brewer.pal(11, "PiYG")
morecols <- colorRampPalette(mypalette)
col.cell <- c("maroon", "green")[class]
heatmap.2(logcpm,col=rev(morecols(50)),trace="none", main="Tumor vs. Normal",ColSideColors=col.cell, scale="row", srtCol=45, cexCol=0.75, cexRow=0.5)
```



#Make a heatmap of your top 100 most variable genes accross all samples 

##1: First find the variance of each gene accross all samples. Hint - you can use the apply function here. var returns the variance ?apply ?var
```{r}
samples <- apply(raw_count_data, 1, var)
```

##2: Next sort the list obtained in step #1 by decreasing variance and use the names function to get the gene names. Subset for the first 100 gene names [1:100]
```{r}
samples <- sort(samples, decreasing = TRUE)
samples1 <- samples[1:100]
geneN<- names(samples1)
```

##3: Now subset your logcpm data selecting for the gene names for the top 100 most variable genes you determined in step #2.
```{r}
New <- logcpm[geneN, ]
```

##4: Check the dimensions of your new data. You should have 100 rows and 6 columns
```{r}
heatmap.2(New,col=rev(morecols(50)),trace="none", main="Tumor vs. Normal",ColSideColors=col.cell, scale="row", srtCol=45, cexCol=0.75, cexRow=0.5)
```



